{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-08-28 ms.tex
Reorganise links to appendices in main text

Below we describe the general approach of the package, then provide
sections outlining potential applications at nested levels of ecological organisation.
Detailed technical documents are provided as supplementary information
(see Appendices); updated versions of these technical documents are
available within the {\plant} package itself.

@@

The default physiological model used in {\plant} largely reflects
that presented in \citet{Falster-2011,Falster-2015}, but with extensions
allowing for diameter growth also to be estimated. We refer to this
default model as \texttt{FFW16}, reflecting the author's initials and
publication year of this paper. The \texttt{FFW16} model estimates rate
of biomass
production for a plant, given its size and the current light
environment. Gross photosynthetic income is calculated from the total
leaf area and the light distribution across the plant's canopy. Costs
of tissue respiration and turnover are then subtracted. The remaining
biomass production is then allocated between growth and reproduction. The key
outputs needed by higher levels of {\plant} are height
growth rate, mortality rate and rate of seed production.  However
additional quantities are also computed, such as total assimilation
and the respiration, turnover and allocation rates for different
tissues, and these intermediates can also be accessed (for further details on the \texttt{FFW16} model see Appendix
\ref{sec:plant}).

While we have implemented a particular physiological model (\texttt{FFW16}), the
package is designed to allow arbitrary additional physiologies to be
introduced. Users are able to modify the physiological model in two
different ways. First, one can vary the parameters of the \texttt{FFW16} model
directly (see Table XXX for list, for examples see Appendix
\ref{sec:parameters}). Second, we allow for changes in key parameters to
bring about changes in other parameters (a hyper-parameterisation), enabling
straightforward modelling of trade-offs. For example, the trait \textsc{lma}
(leaf mass per unit leaf area) is used to estimate the rate of leaf turnover
($k_l$), based on an observed scaling relationship spanning across diverse
vegetation and plant functional types \citep{Wright-2004}:
$$k_l = \beta_{kl1} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{kl2}},$$
where $k_l$ and $\phi$ (lma) are both parameters of the core physiological
model, and $\beta_{kl1}, \beta_{kl2}, \phi_0$ are hyper-parameters. Such
linkages are defined within a user-supplied hyper-parameterisation function
(see Appendix \ref{sec:FFW16} for details). Finally, rather than varying the
parameters of the existing equations in the \texttt{FFW16} model, the
equations themselves can be changed and the higher level functionality of the
package can still be used.  Changed equations would amount to defining a new
physiological model in a new C++ file. As an example, in
the \texttt{FFW16b} physiology, we have modified the function determining
allocation to reproduction.

With a basic physiological model in place, {\plant} can be used to estimate
essential physiological rates for individual plants
(Fig. \ref{fig:plant}). The function \texttt{grow\_plant\_to\_size}
takes a given strategy and light environment and grows the plant,
producing a trajectory of size over time (Fig. \ref{fig:plant}a). This
is achieved by integrating an ordinary differential equation of
size-dependent growth rate (see Appendices \ref{sec:demography} and
\ref{sec:plant} for more details). As the plant grows, allocation to
different tissue types varies and the composition of the plant
changes to include more stem and less leaf
(Fig. \ref{fig:plant}b). This in turn affects maintenance and turnover
costs and the growth rate of the plant (Fig. \ref{fig:plant}c).  This
shift in allocation by itself can generate the widely-observed hump-shaped pattern
of absolute height growth with size \citep{King-2011} and the decline in relative mass growth rate with size from birth onwards \citep{Enquist-2007}.

@@

The extension for handling strong size-asymmetric competition involves
adaptive refining of time points at which new cohorts are introduced
into the population, an idea first applied in \citet{Falster-2011} and
described further in \citet{Falster-2015}.
In the original \textsc{ebt}, cohorts were introduced at a fixed time
interval. Under competition, however, the growth
trajectories of individuals born at similar times can diverge
substantially over time (Fig. \ref{fig:patch}a; Appendix
\ref{sec:demography}, \ref{sec:cohort-spacing} and
\ref{sec:patch}). This can become a problem later during patch
development because much of the population can be located between
widely-spaced cohorts, causing low numerical accuracy. To maintain
numerical accuracy, we use an iterative algorithm to adaptively refine
the cohort spacing until the trajectories of adjacent cohorts remain
adequately resolved throughout development of the patch (see Appendix
\ref{sec:demography} for details). This refinement gives rise to uneven
distribution of cohort introduction times across patch age (Fig.
\ref{fig:patch}a), with considerably tighter spacing of cohorts in
younger patches.

@@

The main numerical challenge in stepping from a single patch to a
metapopulation is to identify the equilibrium seed rain of the
metapopulation. An ``input'' seed rain is required to simulate the
demography of the metapopulation, and this in turn produces an ``output''
number of seeds. Equilibrium occurs when, for each species,
a given input seed rain produces an equivalent output seed rain. Finding this equilibrium
requires finding the root of a multi-dimensional function, subject to
the constraint that this root is stable (the trivial equilibrium
of zero seed rain always exists, but is often not stable; Appendix
\ref{sec:equilibrium}).  Real seed rains are of course quite noisy,
but this approach still allows for the mean tendencies to be estimated.

@@

Second, the metapopulation concept allows for the effects of disturbance
regimes on vegetation structure and dynamics to be properly incorporated. In
{\plant} characteristics for the entire vegetation (metapopulation) -- such
as total leaf area, biomass and productivity -- are obtained by averaging
patch-level outcomes over the frequency-distribution of patch ages (e.g.
Fig. \ref{fig:patch}c).  Similarly, the distributions of  abundance and growth rate
observed in large forest  plots \citep[e.g.][]{Muller-2006, Coomes-2007} arise
as  emergent phenomena. Moreover, the predictions from {\plant}
are for a frequency distribution of patch states across the landscape
(Fig. \ref{fig:emergent}, see more detail in Appendix \ref{sec:emergent}). As such,
size- and patch-structured models provide unique opportunities for linking with
satellite and forest survey data \citep{Moorcroft-2001, Purves-2008}.

@@

With this measure of fitness, we can model trait-based adaptive
dynamics and community assembly of forest communities, following
 \citep{Falster-2015}. This is best thought of as occurring on a fitness landscape, a surface with fitness plotted against one or more species traits
(Fig. \ref{fig:fitness}). The reproductive success of individuals
depends both on their own traits and on densities of other occupants in the current resident community. Fitness landscapes will be positive in regions that allow invasion, and zero
for species that are at equilibrium. The slope of the fitness
landscape with respect to traits indicates the direction of selection
(dashed line in Fig. \ref{fig:fitness}a) indicates selection for
smaller trait values (see Appendix \ref{sec:fitness} for
details)). Fitness landscapes are continually reshaped as species are
introduced and their trait values evolve, reflecting the non-linear demographic feedbacks
that occur under frequency-dependent evolution. Stable coexistence is
possible when species with different traits can mutually invade
\citep{Geritz-1998, Chesson-2000}. Following the fitness gradient may lead to an
evolutionary branching point \citep{Geritz-1998}; at
such a point there is no directional selection (solid line in
Fig. \ref{fig:fitness}a) but invasion is possible both above and below
it in trait space. It is often possible to find a configuration where
fitness is negative everywhere (solid line in Fig. \ref{fig:fitness}b)
except for the zero points where species are present. This community
is immune to invasion from any other strategies and represents a
prediction for the trait mixture that might be favoured by
natural selection \citep{Geritz-1998}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/plant.pdf}
\caption{Physiological dynamics for individual plants varying in size,
  trait and light environment. Solid lines are for the high
  \textsc{lma} (leaf mass per unit area) strategy, dashed lines are
  for the low \textsc{lma} strategy.  (a) Growth trajectories are
  influenced by both light environment and traits.  (b) Over time, the
  fraction of living tissue switches from leaf towards sapwood, with
  high \textsc{lma} species having relatively more mass in leaf than
  low \textsc{lma} species, due to their longer leaf lifespan. (c) Size-dependent growth rates
  (derivative of panel (a)) peak at around 5m high, but location of the peak
  varies with both trait and light level.  (d) Declines in growth rate
  with light vary with plant size and traits.  Whole plant light
  compensation points (indicated by dots) emerge as the point where
  growth rate is zero (x intercept).  The light level used in panels
  (a) and (c) is indicated with coloured ticks.  See Appendix
  \ref{sec:plant} for further details and code.}
\label{fig:plant}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/patch.pdf}
\caption{Population dynamics for two species competing within a patch.
  Red and blue lines are low and high \textsc{lma} species from
  Fig. \ref{fig:plant}.  Lines in panel (a) show growth trajectories
  for individuals on the boundary of successive ``cohorts'' for each
  species, using a schedule of introduction times generated by
  {\plant}'s adaptive algorithm.  The initial growth rate advantage of
  the low \textsc{lma} species (see Figure \ref{fig:plant}c) means it
  quickly overtops the high \textsc{lma} species, suppressing its
  growth.  Self-thinning creates space for the high \textsc{lma}
  species to establish below the canopy of the faster species. Panel
  (b) shows the same trajectories as in panel (a), but with shading
  indicating the density $N$ of individuals at given size and patch
  age (light = low density, dark = high density).  (c) The leaf area
  index at ground level for each species.  The dotted line is the
  total leaf area index (summed across both species), and the lines on
  the right side are averages when integrated over the whole
  metapopulation structure.  See Appendix \ref{sec:patch} for further
  details and code.}
\label{fig:patch}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/emergent.pdf}
\caption{Emergent size-distributions and demography across a metapopulation.
Grey lines show patterns within individual patches differing in time
since disturbance (age), with blue lines indicating patches of a
specific age. Red lines show average relationship across the metapopulation,
obtained by weighting the grey lines by patch and individual abundance.
Panels show (a) patterns of abundance, (b) leaf area and
(c) height growth rate against size. Units of panels (a) and
(b) are amount per unit height per unit ground area.
See Appendix \ref{sec:emergent} for further details and code.}
\label{fig:emergent}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/fitness.pdf}
\caption{Example fitness landscape for \textsc{lma} (leaf mass per
  unit leaf area), showing potential for stable coexistence of
  multiple types.  Fitness is the log of the population growth rate --
  at zero species are at equilibrium, above zero they will increase in
  density.  (a) Landscape generated by a single species (indicated by
  dots).  At \textsc{lma} of 0.1 (the dashed line, open dot) there is
  directional selection for smaller \textsc{lma}.  At \textsc{lma} of
  0.08 (solid line, filled dot) there is a branching point (see inset
  for more detail).  (b) Landscape generated by two species, holding
  the first at \textsc{lma} of 0.08.  Introducing a species at the
  maximum fitness value from panel (a) reduces the region of positive
  fitness considerably (dashed line, open dot).  Dotted lines show
  subsequent invasions and replacements of the second species, the
  solid line shows a stable fitness landscape where both species exist
  on a peak in the fitness landscape.  At this point no further
  invasion is possible.  See Appendix \ref{sec:fitness} for further
  details and code.}
\label{fig:fitness}
\end{figure}

@@

\section{Plant physiological model}\label{sec:FFW16}

See attached file \url{vignettes/physiology.pdf}

@@

See attached file \url{vignettes/demography.pdf}

\section{Plant level properties}\label{sec:plant}

\emph{Note}: due to Manuscript Central not accepting \textsc{html}
files, we have had to convert this (and the following appendices) to
\textsc{pdf}s.  \textsc{html} versions are available on the wiki:
https://github.com/traitecoevo/plant\_paper/wiki --- the \textsc{html}
versions have less pixelated graphics and are easier to copy and
paste out of.

See attached file \url{vignettes/plant.pdf},
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/plant.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/plant}

\section{Cohort spacing algorithm}\label{sec:cohort-spacing}

See attached file \url{vignettes/cohort_spacing.pdf},
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/cohort_spacing.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/cohort_spacing}

\section{Finding demographic equilibrium}\label{sec:equilibrium}

See attached file \url{vignettes/equilibrium.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/equilibrium.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/equilibrium}

\section{Patch level dynamics}\label{sec:patch}

See attached file \url{vignettes/patch.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/patch.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/patch}

\section{Patch level emergent properties}\label{sec:emergent}

See attached file \url{vignettes/emergent.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/emergent.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/emergent}

\section{Calculating fitness}\label{sec:fitness}

See attached file \url{vignettes/fitness.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/fitness.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/fitness}

\section{Modifying parameters of the physiological model}\label{sec:parameters}
See attached file \url{vignettes/parameters.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/parameters.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/parameters}

>>>>

Below we describe the general approach of the package, then provide
sections outlining potential applications at nested levels of ecological organisation.
More detailed technical documents are provided as supplementary information
(see Appendices; updated versions of these technical documents are
available within the {\plant} package itself). These outline \ref{sec:demography}) the system of
equations being solved when modelling demography of plants, patches and metapopulations; \ref{sec:FFW16}) a description of the \texttt{FFW16}
physiological model, and \ref{sec:code}) a compendium of worked code examples.

@@

The default physiological model used in {\plant} largely reflects
that presented in \citet{Falster-2011,Falster-2015}, but with extensions
allowing for diameter growth also to be estimated. We refer to this
default model as \texttt{FFW16}, reflecting the author's initials and
publication year of this paper. The \texttt{FFW16} model estimates rate
of biomass
production for a plant, given its size and the current light
environment. Gross photosynthetic income is calculated from the total
leaf area and the light distribution across the plant's canopy. Costs
of tissue respiration and turnover are then subtracted. The remaining
biomass production is then allocated between growth and reproduction. The key
outputs needed by higher levels of {\plant} are height
growth rate, mortality rate and rate of seed production.  However
additional quantities are also computed, such as total assimilation
and the respiration, turnover and allocation rates for different
tissues, and these intermediates can also be accessed (for further details on the \texttt{FFW16} model see Appendix
\ref{sec:code}).

While we have implemented a particular physiological model (\texttt{FFW16}), the
package is designed to allow arbitrary additional physiologies to be
introduced. Users are able to modify the physiological model in two
different ways. First, one can vary the parameters of the \texttt{FFW16} model
directly (see Table XXX for list, for examples see Appendix
\ref{sec:code}). Second, we allow for changes in key parameters to
bring about changes in other parameters (a hyper-parameterisation), enabling
straightforward modelling of trade-offs. For example, the trait \textsc{lma}
(leaf mass per unit leaf area) is used to estimate the rate of leaf turnover
($k_l$), based on an observed scaling relationship spanning across diverse
vegetation and plant functional types \citep{Wright-2004}:
$$k_l = \beta_{kl1} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{kl2}},$$
where $k_l$ and $\phi$ (lma) are both parameters of the core physiological
model, and $\beta_{kl1}, \beta_{kl2}, \phi_0$ are hyper-parameters. Such
linkages are defined within a user-supplied hyper-parameterisation function
(see Appendix \ref{sec:FFW16} for details). Finally, rather than varying the
parameters of the existing equations in the \texttt{FFW16} model, the
equations themselves can be changed and the higher level functionality of the
package can still be used.  Changed equations would amount to defining a new
physiological model in a new C++ file. As an example, in
the \texttt{FFW16b} physiology, we have modified the function determining
allocation to reproduction.

With a basic physiological model in place, {\plant} can be used to estimate
essential physiological rates for individual plants
(Fig. \ref{fig:plant}). The function \texttt{grow\_plant\_to\_size}
takes a given strategy and light environment and grows the plant,
producing a trajectory of size over time (Fig. \ref{fig:plant}a). This
is achieved by integrating an ordinary differential equation of
size-dependent growth rate (see Appendices \ref{sec:demography} and
\ref{sec:code} for more details). As the plant grows, allocation to
different tissue types varies and the composition of the plant
changes to include more stem and less leaf
(Fig. \ref{fig:plant}b). This in turn affects maintenance and turnover
costs and the growth rate of the plant (Fig. \ref{fig:plant}c).  This
shift in allocation by itself can generate the widely-observed hump-shaped pattern
of absolute height growth with size \citep{King-2011} and the decline in relative mass growth rate with size from birth onwards \citep{Enquist-2007}.

@@

The extension for handling strong size-asymmetric competition involves
adaptive refining of time points at which new cohorts are introduced
into the population, an idea first applied in \citet{Falster-2011} and
described further in \citet{Falster-2015}.
In the original \textsc{ebt}, cohorts were introduced at a fixed time
interval. Under competition, however, the growth
trajectories of individuals born at similar times can diverge
substantially over time (Fig. \ref{fig:patch}a; Appendix
\ref{sec:demography}, and \ref{sec:code}). This can become a problem later during patch
development because much of the population can be located between
widely-spaced cohorts, causing low numerical accuracy. To maintain
numerical accuracy, we use an iterative algorithm to adaptively refine
the cohort spacing until the trajectories of adjacent cohorts remain
adequately resolved throughout development of the patch (see Appendix
\ref{sec:demography} for details). This refinement gives rise to uneven
distribution of cohort introduction times across patch age (Fig.
\ref{fig:patch}a), with considerably tighter spacing of cohorts in
younger patches.

@@

The main numerical challenge in stepping from a single patch to a
metapopulation is to identify the equilibrium seed rain of the
metapopulation. An ``input'' seed rain is required to simulate the
demography of the metapopulation, and this in turn produces an ``output''
number of seeds. Equilibrium occurs when, for each species,
a given input seed rain produces an equivalent output seed rain. Finding this equilibrium
requires finding the root of a multi-dimensional function, subject to
the constraint that this root is stable (the trivial equilibrium
of zero seed rain always exists, but is often not stable; Appendix
\ref{sec:code}).  Real seed rains are of course quite noisy,
but this approach still allows for the mean tendencies to be estimated.

@@

Second, the metapopulation concept allows for the effects of disturbance
regimes on vegetation structure and dynamics to be properly incorporated. In
{\plant} characteristics for the entire vegetation (metapopulation) -- such
as total leaf area, biomass and productivity -- are obtained by averaging
patch-level outcomes over the frequency-distribution of patch ages (e.g.
Fig. \ref{fig:patch}c).  Similarly, the distributions of  abundance and growth rate
observed in large forest  plots \citep[e.g.][]{Muller-2006, Coomes-2007} arise
as  emergent phenomena. Moreover, the predictions from {\plant}
are for a frequency distribution of patch states across the landscape
(Fig. \ref{fig:emergent}, see more detail in Appendix \ref{sec:code}). As such,
size- and patch-structured models provide unique opportunities for linking with
satellite and forest survey data \citep{Moorcroft-2001, Purves-2008}.

@@

With this measure of fitness, we can model trait-based adaptive
dynamics and community assembly of forest communities, following
 \citep{Falster-2015}. This is best thought of as occurring on a fitness landscape, a surface with fitness plotted against one or more species traits
(Fig. \ref{fig:fitness}). The reproductive success of individuals
depends both on their own traits and on densities of other occupants in the current resident community. Fitness landscapes will be positive in regions that allow invasion, and zero
for species that are at equilibrium. The slope of the fitness
landscape with respect to traits indicates the direction of selection
(dashed line in Fig. \ref{fig:fitness}a) indicates selection for
smaller trait values (see Appendix \ref{sec:code} for
details)). Fitness landscapes are continually reshaped as species are
introduced and their trait values evolve, reflecting the non-linear demographic feedbacks
that occur under frequency-dependent evolution. Stable coexistence is
possible when species with different traits can mutually invade
\citep{Geritz-1998, Chesson-2000}. Following the fitness gradient may lead to an
evolutionary branching point \citep{Geritz-1998}; at
such a point there is no directional selection (solid line in
Fig. \ref{fig:fitness}a) but invasion is possible both above and below
it in trait space. It is often possible to find a configuration where
fitness is negative everywhere (solid line in Fig. \ref{fig:fitness}b)
except for the zero points where species are present. This community
is immune to invasion from any other strategies and represents a
prediction for the trait mixture that might be favoured by
natural selection \citep{Geritz-1998}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/plant.pdf}
\caption{Physiological dynamics for individual plants varying in size,
  trait and light environment. Solid lines are for the high
  \textsc{lma} (leaf mass per unit area) strategy, dashed lines are
  for the low \textsc{lma} strategy.  (a) Growth trajectories are
  influenced by both light environment and traits.  (b) Over time, the
  fraction of living tissue switches from leaf towards sapwood, with
  high \textsc{lma} species having relatively more mass in leaf than
  low \textsc{lma} species, due to their longer leaf lifespan. (c) Size-dependent growth rates
  (derivative of panel (a)) peak at around 5m high, but location of the peak
  varies with both trait and light level.  (d) Declines in growth rate
  with light vary with plant size and traits.  Whole plant light
  compensation points (indicated by dots) emerge as the point where
  growth rate is zero (x intercept).  The light level used in panels
  (a) and (c) is indicated with coloured ticks.  See Appendix
  \ref{sec:code} for further details and code.}
\label{fig:plant}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/patch.pdf}
\caption{Population dynamics for two species competing within a patch.
  Red and blue lines are low and high \textsc{lma} species from
  Fig. \ref{fig:plant}.  Lines in panel (a) show growth trajectories
  for individuals on the boundary of successive ``cohorts'' for each
  species, using a schedule of introduction times generated by
  {\plant}'s adaptive algorithm.  The initial growth rate advantage of
  the low \textsc{lma} species (see Figure \ref{fig:plant}c) means it
  quickly overtops the high \textsc{lma} species, suppressing its
  growth.  Self-thinning creates space for the high \textsc{lma}
  species to establish below the canopy of the faster species. Panel
  (b) shows the same trajectories as in panel (a), but with shading
  indicating the density $N$ of individuals at given size and patch
  age (light = low density, dark = high density).  (c) The leaf area
  index at ground level for each species.  The dotted line is the
  total leaf area index (summed across both species), and the lines on
  the right side are averages when integrated over the whole
  metapopulation structure.  See Appendix \ref{sec:code} for further
  details and code.}
\label{fig:patch}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/emergent.pdf}
\caption{Emergent size-distributions and demography across a metapopulation.
Grey lines show patterns within individual patches differing in time
since disturbance (age), with blue lines indicating patches of a
specific age. Red lines show average relationship across the metapopulation,
obtained by weighting the grey lines by patch and individual abundance.
Panels show (a) patterns of abundance, (b) leaf area and
(c) height growth rate against size. Units of panels (a) and
(b) are amount per unit height per unit ground area.
See Appendix \ref{sec:code} for further details and code.}
\label{fig:emergent}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/fitness.pdf}
\caption{Example fitness landscape for \textsc{lma} (leaf mass per
  unit leaf area), showing potential for stable coexistence of
  multiple types.  Fitness is the log of the population growth rate --
  at zero species are at equilibrium, above zero they will increase in
  density.  (a) Landscape generated by a single species (indicated by
  dots).  At \textsc{lma} of 0.1 (the dashed line, open dot) there is
  directional selection for smaller \textsc{lma}.  At \textsc{lma} of
  0.08 (solid line, filled dot) there is a branching point (see inset
  for more detail).  (b) Landscape generated by two species, holding
  the first at \textsc{lma} of 0.08.  Introducing a species at the
  maximum fitness value from panel (a) reduces the region of positive
  fitness considerably (dashed line, open dot).  Dotted lines show
  subsequent invasions and replacements of the second species, the
  solid line shows a stable fitness landscape where both species exist
  on a peak in the fitness landscape.  At this point no further
  invasion is possible.  See Appendix \ref{sec:code} for further
  details and code.}
\label{fig:fitness}
\end{figure}

@@

@@

Describes the system of equations being solved in {\plant} when modelling demography of plants, patches and metapopulations. See attached file \url{vignettes/demography.pdf}

\section{Description of the \texttt{FFW16} physiological model}\label{sec:FFW16}

Describes the core physiological model embedded in plant (\texttt{FFW16}). See attached file \url{vignettes/physiology.pdf}

\section{Code examples}\label{sec:code}

A compendium of worked code examples showing how to intearct with {\plant} from R
See attached file \url{vignettes/combined.pdf}.
